|

### 2019-09-13 analysis/paper/manuscript.Rmd
Address comments from reviewer 3 (namely add an appendix listing some examples) and reviewer 2

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2009; @maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) of presenting information at stakeholder meetings, and to fisheries managers, are simply insufficient to convey the richness of the information available. More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community. As a result, an increasing number of scientists across a wide range of disciplines, including fisheries research, are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018].

@@

The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018]. The majority of this sequence of events, sometimes called "the data pipeline" [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis. However, decisions made along the data pipeline increasingly influence the outcome of the study. Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis. Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible. A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015]. Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009].

The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (e.g. _________). We  believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued. While these tools may not reveal the entire data-pipeline, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011]. In the case of NCAM, considerable statistical experience and expertise is required simply to run the model. However, the dashboard removes many of these difficulties. The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document that can be distributed prior to meetings and, as such, participants can assess the results and diagnostics at their leisure. Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.

@@

We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.

>>>>

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) of presenting information at stakeholder meetings, and to fisheries managers, are simply insufficient to convey the richness of the information available. More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

@@

The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018]. The majority of this sequence of events, sometimes called "the data pipeline" [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis. However, decisions made along the data pipeline increasingly influence the outcome of the study. Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis. Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible. A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015]. Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009; @lowndes2017].

The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (see Appendix A). We  believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued. While these tools may not reveal the entire data-pipeline, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011]. In the case of NCAM, considerable statistical experience and expertise is required simply to run the model. However, the dashboard removes many of these difficulties. The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document that can be distributed prior to meetings and, as such, participants can assess the results and diagnostics at their leisure. Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.

@@

We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful. We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and two anonymous reviewers.

# Appendix A

Partial list of applications developed to explore and communicate marine and fisheries science:

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"
